---
title: "de_analysis_wt_ko"
author: "Yunus Ozekin"
date: "2025-03-04"
output: html_document
---


```{css, echo = F}
caption {
      color: blue;
      font-weight: bold;
      font-size: 2.0em;
    }
```


```{r, results='hide', message=FALSE, warning=FALSE, error=FALSE}
library(Seurat)
library(sctransform)
library(ggplot2)
library(dplyr)
library(ggsci)
library(cowplot)
library(patchwork)
library(RColorBrewer)
library(tidyverse)
library(forcats)
library(tidyr)
library(ComplexHeatmap)
options(future.globals.maxSize = 4000 * 1024^2)
mem.maxVSize(vsize = 32000000000)
```

```{r}
mypal <- c("#68023F","#008169","#EF0096","#00DCB5","#FFCFE2",
           "#003C86","#9400E6","#009FFA","#FF71FD","#7CFFFA",
           "#6A0213","#008607","#F60239","#00E307","#FFDC3D","#ffea8a","#001310","#ffe9f1")

#mypal2 <- c("#5A5156", "#E4E1E3", "#F6222E", "#FE00FA", "#16FF32", "#3283FE", "#FEAF16", "#B00068", "#1CFFCE", "#90AD1C", "#2ED9FF", "#DEA0FD", "#AA0DFE", "#F8A19F", "#325A9B", "#C4451C", "#1C8356", "#85660D", "#B10DA1", "#FBE426", "#1CBE4F", "#FA0087", "#FC1CBF", "#F7E1A0", "#C075A6", "#782AB6", "#AAF400", "#BDCDFF", "#822E1C", "#B5EFB5", "#7ED7D1", "#1C7F93", "#D85FF7", "#683B79", "#66B0FF", "#3B00FB")

```

# Read Data

```{r, echo=FALSE}
int.seurat <- readRDS(file = "RDSfiles/All.integrated.clustered.RDS")
all.markers <- readRDS(file = "RDSfiles/All.markers.int5.RDS")
```

# Remove XY genes
```{r}
library(biomaRt)

# Connect to Ensembl database
mart <- useMart("ensembl", dataset = "mmusculus_gene_ensembl")

# Get genes on chromosomes X and Y
xy_genes <- getBM(attributes = c("external_gene_name", "chromosome_name"), 
                   filters = "chromosome_name", 
                   values = c("X", "Y"), 
                   mart = mart)$external_gene_name

int.seurat <- subset(int.seurat, features = setdiff(rownames(int.seurat), xy_genes))

any(rownames(int.seurat) %in% xy_genes)  # Should return FALSE

```

# Remove RPL and RPS genes and Kcnj2
```{r}
library(biomaRt)

# Connect to Ensembl database
mart <- useMart("ensembl", dataset = "mmusculus_gene_ensembl")

# Get rRNA genes
rRNA_genes <- getBM(
  attributes = c("external_gene_name"),
  filters = "biotype",
  values = "rRNA",
  mart = mart)$external_gene_name

int.seurat <- subset(int.seurat, features = setdiff(rownames(int.seurat), rRNA_genes))

# Get Rpl and Rps genes
rp_genes <- grep("^Rps|^Rpl", rownames(int.seurat), value = TRUE)
int.seurat <- subset(int.seurat, features = setdiff(rownames(int.seurat), rp_genes))

any(rownames(int.seurat) %in% rRNA_genes)  # Should return FALSE
any(rownames(int.seurat) %in% rp_genes)  # Should return FALSE


int.seurat <- subset(int.seurat, features = setdiff(rownames(int.seurat), "Kcnj2"))
```

```{r}
library(biomaRt)

# Connect to Ensembl database
mart <- useMart("ensembl", dataset = "mmusculus_gene_ensembl")

# Get Gm genes
gm_genes <- grep("^Gm", rownames(int.seurat), value = TRUE)
int.seurat <- subset(int.seurat, features = setdiff(rownames(int.seurat), gm_genes))

any(rownames(int.seurat) %in% gm_genes)  # Should return FALSE
```

# Remove MT genes
```{r}
library(biomaRt)

# Connect to Ensembl database
mart <- useMart("ensembl", dataset = "mmusculus_gene_ensembl")

# Get MT genes
MT_genes <- getBM(attributes = c("external_gene_name", "chromosome_name"), 
                   filters = "chromosome_name", 
                   values = c("MT"), 
                   mart = mart)$external_gene_name

int.seurat <- subset(int.seurat, features = setdiff(rownames(int.seurat), MT_genes))

any(rownames(int.seurat) %in% MT_genes)  # Should return FALSE

```

# Proportion of cells in each cluster across each sample
```{r, echo=FALSE}

DefaultAssay(int.seurat) <- "SCT"
Idents(int.seurat) <- "orig.ident"
#int.seurat <- subset(x = int.seurat, idents = 'HET', invert = TRUE)

pt <- table(int.seurat$Clusters, int.seurat$KO_type)
pt <- as.data.frame(pt)

pt <- filter(pt, Var2 != "HET")

pt$Total <- ifelse(grepl("KO", pt$Var2), 11658, 22238)
pt$Percentage <- ((pt$Freq)/(pt$Total))*100


KO_type <- c("WT", "KO")

#png(filename = "WT_KO_Proportions.png", width = 5, height = 5, units = "in", res = 300)
ggplot(pt, aes(x = fct_relevel(Var2, "WT"), y = Freq, fill = Var1)) +
  theme_bw(base_size = 15) +
  geom_col(position = "fill", width = 0.5) +
  xlab("Sample") +
  ylab("Proportion") +
  theme(legend.title = element_blank()) + 
  scale_fill_manual(values = mypal)
#dev.off()

pt_table <- subset (pt, select = -c(Var2, Freq, Total)) %>% 
  pivot_wider(names_from = Var1, values_from = Percentage)
pt_table

```
# Remove Het from analysis
```{r}
unique(Idents(int.seurat))
int.seurat <- subset(x = int.seurat, idents = c("Kir2KO_3"), invert = TRUE)
unique(Idents(int.seurat))
```


```{r}
DefaultAssay(int.seurat) <- "SCT"
Idents(int.seurat) <- "Clusters"
int.seurat$celltype.ko <- paste(Idents(object = int.seurat), int.seurat$KO_type, sep = '.') 
Idents(int.seurat) <- "celltype.ko"
levels(int.seurat) <- c("Mes_1.KO", "Mes_1.WT", "Mes_2.KO", "Mes_2.WT", "Mes_3.KO", "Mes_3.WT", "Mes_4.KO", "Mes_4.WT", "Prolif_1.KO", "Prolif_1.WT", "Prolif_2.KO",  "Prolif_2.WT", "Prolif_3.KO", "Prolif_3.WT", "Ep_1.KO", "Ep_1.WT", "Ep_2.KO", "Ep_2.WT", "Imm.KO", "Imm.WT", "Schw.KO", "Schw.WT", "End.KO", "End.WT")

```

# UMAP by Cluster, KO_type, and DefiningName
```{r, echo=FALSE}
png(filename = "DimPlot_Clusters_WT_KO.png", width = 7.5, height = 5, units = "in", res = 300)
DimPlot(int.seurat, group.by = "Clusters", cols = mypal, label = T, repel = T)
dev.off()
png(filename = "DimPlot_WT_KO.png", width = 7.5, height = 5, units = "in", res = 300)
DimPlot(int.seurat, group.by = "KO_type", label = T, repel = T) + scale_color_igv()
dev.off()
```
```{r, fig.width=12}
#p1 <- DimPlot(int.seurat, group.by = "Clusters", cols = mypal, label = T, repel = T)
#p2 <-FeaturePlot(int.seurat, features = c("Osr2", "Runx2", "Msx1", "Epha3", "Kcnj2"))
#Idents(int.seurat) <- "Clusters"
#png(filename = "int_marker_dot.png", width = 12, height = 8, units = "in", res = 300)
#p1 + p2
#DotPlot(int.seurat, features = c("Osr2", "Shox2", "Runx2", "Msx1", "Ogn", "Mecom", "Epha3", "Egfl6", "Igfbp2", "Trp63", "Cldn7", "Krt18", "Kcnj2")) + RotatedAxis()
#dev.off()
```
# Mes1
```{r, echo=FALSE, fig.width=9, fig.height=4}
Mes1_de_exp <- FindMarkers(int.seurat, ident.1 = "Mes_1.WT", ident.2 = "Mes_1.KO", logfc.threshold = 0.5, verbose = FALSE)
head(Mes1_de_exp, n = 30)

DotPlot(int.seurat, idents = c("Mes_1.WT", "Mes_1.KO"), features = rownames(head(Mes1_de_exp, n = 30)), scale = F) + RotatedAxis() + labs(title = "Mes_1 DE Genes")
```

# Mes2
```{r, echo=FALSE, fig.width=9, fig.height=4}
Mes2_de_exp <- FindMarkers(int.seurat, ident.1 = "Mes_2.WT", ident.2 = "Mes_2.KO", logfc.threshold = 0.5, verbose = FALSE)
head(Mes2_de_exp, n = 30)


DotPlot(int.seurat, idents = c("Mes_2.WT", "Mes_2.KO"), features = rownames(head(Mes2_de_exp, n = 30)), scale = F) + RotatedAxis() + labs(title = "Mes_2 DE Genes")
```

# Mes3
```{r, echo=FALSE, fig.width=9, fig.height=4}
Mes3_de_exp <- FindMarkers(int.seurat, ident.1 = "Mes_3.WT", ident.2 = "Mes_3.KO", logfc.threshold = 0.5, verbose = FALSE)
head(Mes3_de_exp, n = 30)


DotPlot(int.seurat, idents = c("Mes_3.WT", "Mes_3.KO"), features = rownames(head(Mes3_de_exp, n = 30)), scale = F) + RotatedAxis() + labs(title = "Mes_3 DE Genes")
```

# Mes4
```{r, echo=FALSE, fig.width=9, fig.height=4}
Mes4_de_exp <- FindMarkers(int.seurat, ident.1 = "Mes_4.WT", ident.2 = "Mes_4.KO", logfc.threshold = 0.5, verbose = FALSE)
head(Mes4_de_exp, n = 30)


DotPlot(int.seurat, idents = c("Mes_4.WT", "Mes_4.KO"), features = rownames(head(Mes4_de_exp, n = 30)), scale = F) + RotatedAxis() + labs(title = "Mes_4 DE Genes")

```

# Ectoderm 1
```{r, echo=FALSE, fig.width=9, fig.height=4}
Ep1_de_exp <- FindMarkers(int.seurat, ident.1 = "Ep_1.WT", ident.2 = "Ep_1.KO", logfc.threshold = 0.5, verbose = FALSE)
head(Ep1_de_exp, n = 30)

DotPlot(int.seurat, idents = c("Ep_1.WT", "Ep_1.KO"), features = rownames(head(Ep1_de_exp, n = 30)), scale = F) + RotatedAxis() + labs(title = "Ep_1 DE Genes")
```

# Ectoderm 2
```{r, echo=FALSE, fig.width=9, fig.height=4}
Ep2_de_exp <- FindMarkers(int.seurat, ident.1 = "Ep_2.WT", ident.2 = "Ep_2.KO", logfc.threshold = 0.5, verbose = FALSE)
head(Ep2_de_exp, n = 30)

DotPlot(int.seurat, idents = c("Ep_2.WT", "Ep_2.KO"), features = rownames(head(Ep2_de_exp, n = 30)), scale = F) + RotatedAxis() + labs(title = "Ep_2 DE Genes")
```

# Proliferating 1
```{r, echo=FALSE, fig.width=9, fig.height=4}
Prolif1_de_exp <- FindMarkers(int.seurat, ident.1 = "Prolif_1.WT", ident.2 = "Prolif_1.KO", logfc.threshold = 0.5, verbose = FALSE)
head(Prolif1_de_exp, n = 30)


DotPlot(int.seurat, idents = c("Prolif_1.WT", "Prolif_1.KO"), features = rownames(head(Prolif1_de_exp, n = 30)), scale = F) + RotatedAxis() + labs(title = "Prolif_1 DE Genes")
```

# Proliferating 2
```{r, echo=FALSE, fig.width=9, fig.height=4}
Prolif2_de_exp <- FindMarkers(int.seurat, ident.1 = "Prolif_2.WT", ident.2 = "Prolif_2.KO", logfc.threshold = 0.5, verbose = FALSE)
head(Prolif2_de_exp, n = 30)

DotPlot(int.seurat, idents = c("Prolif_2.WT", "Prolif_2.KO"), features = rownames(head(Prolif2_de_exp, n = 30)), scale = F) + RotatedAxis() + labs(title = "Prolif_2 DE Genes")
```

# Proliferating 3
```{r, echo=FALSE, fig.width=9, fig.height=4}
Prolif3_de_exp <- FindMarkers(int.seurat, ident.1 = "Prolif_3.WT", ident.2 = "Prolif_3.KO", logfc.threshold = 0.5, verbose = FALSE)
head(Prolif3_de_exp, n = 30)

DotPlot(int.seurat, idents = c("Prolif_3.WT", "Prolif_3.KO"), features = rownames(head(Prolif3_de_exp, n = 30)), scale = F) + RotatedAxis() + labs(title = "Prolif_3 DE Genes")
```
```{r}
write.csv(Ep1_de_exp,"MarkerGenes/Ep1.csv", row.names = TRUE)
write.csv(Ep2_de_exp,"MarkerGenes/Ep2.csv", row.names = TRUE)
write.csv(Mes1_de_exp,"MarkerGenes/Mes1.csv", row.names = TRUE)
write.csv(Mes2_de_exp,"MarkerGenes/Mes2.csv", row.names = TRUE)
write.csv(Mes3_de_exp,"MarkerGenes/Mes3.csv", row.names = TRUE)
write.csv(Mes4_de_exp,"MarkerGenes/Mes4.csv", row.names = TRUE)
write.csv(Prolif1_de_exp,"MarkerGenes/Prolif1.csv", row.names = TRUE)
write.csv(Prolif2_de_exp,"MarkerGenes/Prolif2.csv", row.names = TRUE)
write.csv(Prolif3_de_exp,"MarkerGenes/Prolif3.csv", row.names = TRUE)

```


```{r}
library(dplyr)
library(gt)

# Prepare data
top_genes <- Mes4_de_exp %>%
  tibble::rownames_to_column("gene") %>%
  arrange(p_val_adj) %>%
  select(gene, avg_log2FC, pct.1, pct.2, p_val_adj) %>%
  head(30)

# Create compact gt table
gt_table <- gt(top_genes) %>%
  fmt_number(columns = c(avg_log2FC, pct.1, pct.2), decimals = 3) %>%
  cols_label(
    gene = "Gene",
    avg_log2FC = "log2FC",
    pct.1 = "Pct WT",
    pct.2 = "Pct KO",
    p_val_adj = "Adj P-Value"
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = list(
      cells_column_labels(everything()),
      cells_title(groups = "title")
    )
  ) %>%
  tab_header(title = "Top 30 Significant Genes Mes_4") %>%
  tab_options(
    table.width = px(400),
    column_labels.font.size = px(12),
    table.font.size = px(12)
  )

# Save the table as a PNG
gtsave(gt_table, "top_genes_table_Mes_4.png")
```

# Generate lists of genes of interest  

```{r}
#Bmp ligands
bmp.ligands <- c("Bmp1", "Bmp2", "Bmp3", "Bmp4", "Bmp5", "Bmp6", "Bmp7", "Bmp8a", "Bmp8b", "Bmp11", "Bmp15", "Tgfb1", "Tgfb2", "Tgfb3")

#Bmp receptors
bmp.receptors <- c("Bmpr1a", "Bmpr1b", "Bmpr2", "Tgfbr1", "Tgfbr2", "Tgfbr3")

#smads
smads <- c("Smad1", "Smad2", "Smad3", "Smad4", "Smad5", "Smad6", "Smad7")

#Bmp inhibitors
bmp.inhibitors <- c("Bambi", "Chrd", "Grem1", "Dcn", "Fst", "Ltpb1", "Ltbp2", "Nog", "Tgfbi", "Tgfbrap1", "Thbs1", "Sostdc1")

#Wnt ligands
wnt.ligands <- c("Wnt1", "Wnt2", "Wnt2b", "Wnt3", "Wnt3a", "Wnt4", "Wnt5a", "Wnt5b", "Wnt6", "Wnt7a", "Wnt7b", "Wnt8a", "Wnt8b", "Wnt9a", "wnt9b", "Wnt10a", "Wnt10b", "Wnt11", "Wnt16b")

#Wnt receptors
wnt.receptors <- c("Frzd1", "Frzd2", "Frzd3", "Frzd4", "Frzd5", "Frzd6", "Frzd7", "Frzd8", "Frzd9", "Frzd10", "Lrp1", "Lrp5", "Lrp6", "Lrp3", "Lrp8", "Lrp10", "Lrp11", "Lrpap1", "Musk")

#Wnt inhibitors
wnt.inhibitors <- c("Dkk1", "Dkk2", "Dkk3", "Sfrp1", "Sfrp2", "Sfrp5", "Wif1", "Wise", "Sostdc1", "Cerl1", "Igfb4", "Shisa", "Shisa3", "Shisa4", "Shisa5", "Rspo1", "Rspo2", "Rspo3", "Norrin")

#hh ligands
hh.ligands <- c("Shh", "Ihh", "Dhh")

#hh receptors
hh.receptors <- c("Smo", "Ptch1", "Ptch2")

#Fgf ligands
fgf.ligands <- c("Fgf1", "Fgf2", "Fgf3", "Fgf4", "Fgf5", "Fgf6", "Fgf7", "Fgf8", "Fgf9", "Fgf10", "Fgf18", "Fgf23")

#Fgf receptors
fgf.receptors <- c("Fgfr1", "Fgfr2", "Fgfr2b", "Fgfr3", "Fgfr4")

#Igf ligands
igf.ligands <- c("Igf1", "Igf2")

#Igf receptors
igf.receptors <- c("Igf1r", "Igflr1", "Insr")

#Pdgf ligands
pdgf.ligands <- c("Pdgfa", "Pdgfb", "Pdgfc", "Pdgfd")

#Pdgf receptors
pdgf.receptors <- c("Pdgfra", "Pdgfrb", "Pdgfrl")

#Egf ligands
egf.ligands <- c("Egf", "Tgfa", "Hbegf","Btc", "Areg", "Ereg", "Epgn")

#Egf receptors
egf.receptors <- c("Egfr", "Erbb1", "Her1")

#Notch ligands
notch.ligands <- c("Jag1", "Jag2", "Dll1", "Dll3", "Dll4")

#Notch receptors
notch.receptors <- c("Notch1", "Notch2", "Notch3", "Notch4")

#Notch regulators
notch.regulators <- c("Lfng", "Mfng", "Rfng")

#Notch targets
notch.targets <- c("Hes1", "Hes2", "Hes3", "Hes4", "Hes5", "Hes6", "Hes7", "Hey1", "Hey2", "Heyl")

#Endothelin ligands
Edn.ligands <- c("Edn1", "Edn2", "Edn3")

#Endothelin receptors
Edn.receptors <- c("Ednra", "Ednrb")

#Endothelin regulators
Edn.regulators <- c("Ece1", "Ecel1", "Ece2")

int.seurat.subset <- subset(int.seurat, idents = c("Mes_1.KO", "Mes_1.WT", "Mes_2.KO", "Mes_2.WT", "Mes_3.KO", "Mes_3.WT", "Mes_4.KO", "Mes_4.WT"))


```

# BMP ligands, receptors, smads, and inhibitors
```{r, echo=FALSE, warning=FALSE}
DefaultAssay(int.seurat.subset) = "SCT"

#png(filename = "BMP_Genes.png", width = 8.5, height = 5, units = "in", res = 300)
DotPlot(int.seurat.subset, features = c(bmp.ligands, bmp.receptors, smads)) + RotatedAxis() + ggtitle(label = "Bmp") + theme(axis.text.x = element_text(size = 15), axis.text.y = element_text(size = 15), axis.title = element_blank())
#dev.off()
```

# Notch ligands, receptors, regulators, and targets
```{r, echo=FALSE, warning=FALSE}
#png(filename = "Notch_Genes.png", width = 8, height = 5, units = "in", res = 300)
DotPlot(int.seurat.subset, features = c(notch.ligands, notch.receptors)) + RotatedAxis() + ggtitle(label = "Notch") + theme(axis.text.x = element_text(size = 15), axis.text.y = element_text(size = 15), axis.title = element_blank())
#dev.off()
```

# Wnt ligands, receptors, and inhbitors
```{r, echo=FALSE, warning=FALSE}
#png(filename = "Wnt_Genes.png", width = 8, height = 5, units = "in", res = 300)
DotPlot(int.seurat.subset, features = c(wnt.ligands, wnt.receptors)) + RotatedAxis() + ggtitle(label = "Wnt")  + theme(axis.text.x = element_text(size = 15), axis.text.y = element_text(size = 15), axis.title = element_blank())
#dev.off()
```

# Hh ligands and receptors
No changes to hh ligands or receptors  
```{r, echo=FALSE, warning=FALSE}
#png(filename = "Hh_Genes.png", width = 8, height = 5, units = "in", res = 300)
DotPlot(int.seurat.subset, features = c(hh.ligands, hh.receptors)) + RotatedAxis() + ggtitle(label = "Hh")  + theme(axis.text.x = element_text(size = 15), axis.text.y = element_text(size = 15), axis.title = element_blank())
#dev.off()
```

# Fgf ligands and receptors
```{r, echo=FALSE, warning=FALSE}
#png(filename = "fgf_Genes.png", width = 8, height = 5, units = "in", res = 300)
DotPlot(int.seurat.subset, features = c(fgf.ligands, fgf.receptors)) + RotatedAxis() + ggtitle(label = "Fgf")  + theme(axis.text.x = element_text(size = 15), axis.text.y = element_text(size = 15), axis.title = element_blank())
#dev.off()
```

# Igf ligands and receptors
```{r, echo=FALSE, warning=FALSE}
#png(filename = "igf_Genes.png", width = 8, height = 5, units = "in", res = 300)
DotPlot(int.seurat.subset, features = c(igf.ligands, igf.receptors)) + RotatedAxis() + ggtitle(label = "Igf")  + theme(axis.text.x = element_text(size = 15), axis.text.y = element_text(size = 15), axis.title = element_blank())
#dev.off()
```

# Egf ligands and receptors
```{r, echo=FALSE, warning=FALSE}
#png(filename = "egf_Genes.png", width = 8, height = 5, units = "in", res = 300)
DotPlot(int.seurat.subset, features = c(egf.ligands, egf.receptors)) + RotatedAxis() + ggtitle(label = "Egf")  + theme(axis.text.x = element_text(size = 15), axis.text.y = element_text(size = 15), axis.title = element_blank())
#dev.off()
```

# Pdgf ligands and receptors
```{r, echo=FALSE, warning=FALSE}
#png(filename = "Pdgf_Genes.png", width = 8, height = 5, units = "in", res = 300)
DotPlot(int.seurat.subset, features = c(pdgf.ligands, pdgf.receptors)) + RotatedAxis() + ggtitle(label = "Pdgf")  + theme(axis.text.x = element_text(size = 15), axis.text.y = element_text(size = 15), axis.title = element_blank())
#dev.off()
```

# Edn ligands and receptors
```{r, echo=FALSE, warning=FALSE}
#png(filename = "Edn_Genes.png", width = 8, height = 5, units = "in", res = 300)
DotPlot(int.seurat.subset, features = c(Edn.ligands, Edn.receptors, Edn.regulators)) + RotatedAxis() + ggtitle(label = "Edn")  + theme(axis.text.x = element_text(size = 15), axis.text.y = element_text(size = 15), axis.title = element_blank())
#dev.off()
```

```{r}
#png(filename = "Genes_for_Emily.png", width = 8, height = 5, units = "in", res = 300)
#DotPlot(int.seurat.subset, features = c("Fos", "Egr1", "Fn1", "Jun", "Nr4a1")) + RotatedAxis() + theme(axis.text.x = element_text(size = 15), axis.text.y = element_text(size = 15), axis.title = element_blank())
#dev.off()
```

```{r}
int.seurat@meta.data$Clusters = factor(int.seurat@meta.data$Clusters, levels = c("Mes_1", "Mes_2", "Mes_3", "Mes_4", "Prolif_1", "Prolif_2", "Prolif_3", "Ep_1", "Ep_2", "End", "Imm", "Schw"))
int.seurat@meta.data$KO_type = factor(int.seurat@meta.data$KO_type, levels = c("WT", "KO"))

#png(filename = "BMP_pathway_genes.png", width = 7, height = 7, units = "in", res = 300)
VlnPlot(
  object = int.seurat,
  features = c("Smad1", "Shox2", "Barx1", "Msx1", "Wnt5a", "Satb2", "Fgf10", "Runx1", "Runx2", "Runx3", "Shh"),
  group.by = "Clusters",
  split.by = "KO_type",
  pt.size = 0,
  stack = TRUE,
  flip = TRUE,
  cols = c("Purple", "Green")
) + 
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text.y = element_text(face = "bold", angle = 0),
    panel.grid.major = element_blank()
  )
#dev.off()
```

```{r}
sessionInfo()
```