---
title: "int5 Mes subset DoRothEA R Notebook"
output: 
  html_notebook:
    toc: TRUE
    toc_depth: 3
    toc_float: TRUE
---

```{css, echo = F}
caption {
      color: blue;
      font-weight: bold;
      font-size: 2.0em;
    }
```

# libraries
```{r, warning=F}
suppressPackageStartupMessages({
  library(dorothea)
  library(dplyr)
  library(Seurat)
  library(tibble)
  library(pheatmap)
  library(tidyr)
  library(viper)
  library(ggsci)
  library(ggplot2)
})
```

# Read data
```{r}
int5 <- readRDS("RDSfiles/All.integrated.clustered.RDS")

Idents(int5) <- "celltype.ko"


#Reorder Idents to keep things pretty
Idents(int5) <- "Clusters"
levels(int5) <- c("Mes_1","Mes_2","Mes_3","Mes_4","Prolif_1","Prolif_2","Prolif_3","Ep_1","Ep_2","Imm","Schw","End","RBCs")

DimPlot(int5) + scale_color_igv()
```

# memSubset out 
```{r}
int5 <- subset(int5, idents = c("Mes_1","Mes_2","Mes_3","Mes_4"))
Idents(int5) <- "KO_type"
int5 <- subset(int5, idents = c("WT", "KO"))
Idents(int5) <- "KO_type"
DimPlot(int5) + scale_color_igv()
```
```{r}
int5$Clusters <- paste(int5$Clusters, int5$KO_type, sep = "_")
Idents(int5) <- "KO_type"
p1 <- DimPlot(int5, group.by = "Clusters_KOtype", cells = WhichCells(int5, idents = "KO"), label = T, repel = T) + 
  scale_color_igv() + NoLegend() + ggtitle("KO")
p2 <- DimPlot(int5, group.by = "Clusters_KOtype", cells = WhichCells(int5, idents = "WT"), label = T, repel = T) + 
  scale_color_igv() + NoLegend() + ggtitle("WT")
p1 + p2
```


# Run DoRothEA
```{r}
Idents(int5) <- "Clusters_KOtype"

levels(int5) <- c("Mes_1_KO", "Mes_1_WT", "Mes_2_KO", "Mes_2_WT", "Mes_3_KO", "Mes_3_WT", "Mes_4_KO", "Mes_4_WT", "Mes_5_KO", "Mes_5_WT", "Mes_6_KO", "Mes_6_WT", "Mes_7_KO", "Mes_7_WT", "Mes_8_KO", "Mes_8_WT")
```

```{r}
## We read Dorothea Regulons for mouse:
dorothea_regulon_mouse <- get(data("dorothea_mm", package = "dorothea"))

## Skipping the filtration by confidence level
#regulon <- dorothea_regulon_mouse %>%
#    dplyr::filter(confidence %in% c("A","B","C"))

## We compute Viper Scores 
int5 <- run_viper(int5, dorothea_regulon_mouse, assay_key = "SCT",
                  options = list(method = "scale", minsize = 4, 
                                 eset.filter = FALSE, cores = 1, 
                                 verbose = FALSE))
```

```{r}
head(dorothea_regulon_mouse)
dim(dorothea_regulon_mouse)
```

```{r}
## We compute the Nearest Neighbours to perform cluster
DefaultAssay(object = int5) <- "dorothea"
int5 <- ScaleData(int5)
int5 <- RunPCA(int5, features = rownames(int5), verbose = FALSE)
int5 <- FindNeighbors(int5, dims = 1:30, verbose = FALSE)
int5 <- FindClusters(int5, resolution = 0.5, verbose = FALSE)

int5 <- RunUMAP(int5, dims = 1:30, umap.method = "uwot", metric = "cosine")
```
# UMAP based on DoRothEA TF activity
```{r, fig.height=7, fig.width=10}
p1 <- DimPlot(int5) + scale_color_igv()
p2 <- DimPlot(int5, group.by = "Clusters") + scale_color_igv()
p3 <- DimPlot(int5, group.by = "dorothea_snn_res.0.5", split.by = "KO_type") + scale_color_igv() + NoLegend()
(p1 + p2) / p3
```
```{r}
table(int5$dorothea_snn_res.0.5,int5$Clusters_KOtype)
```


```{r}
saveRDS(int5, file = "RDSfiles/int5.Mes.subset.dorothea.RDS")
```

```{r}
int5 <- readRDS(file = "RDSfiles/int5.Mes.subset.dorothea.RDS")
```

# Top10 differntial TF activities per Seurat cluster
```{r}
DefaultAssay(object = int5) <- "dorothea"
Idents(int5) <- "Clusters_KOtype"
All.markers <- FindAllMarkers(int5, only.pos = TRUE, min.pct = 0.25, 
                               logfc.threshold = 0.25, verbose = FALSE)
top10.markers <- All.markers %>% group_by(cluster) %>% top_n(n=-10, wt = p_val)
top10.markers
```
# Top10 TF activities per DoRothEA cluster
```{r}
Idents(int5) <- "dorothea_snn_res.0.5"
All.markers.do <- FindAllMarkers(int5, only.pos = TRUE, min.pct = 0.25, 
                               logfc.threshold = 0.25, verbose = FALSE)
top10.markers.do <- All.markers.do %>% group_by(cluster) %>% top_n(n=-10, wt = p_val)
top10.markers.do
```

# TF activity per cell population

```{r}
## We transform Viper scores, scaled by seurat, into a data frame to better 
## handling the results
viper_scores_df <- GetAssayData(int5, slot = "scale.data", 
                                    assay = "dorothea") %>%
  data.frame(check.names = F) %>%
  t()

## We create a data frame containing the cells and their clusters
CellsClusters <- data.frame(cell = names(Idents(int5)), 
                            cell_type = as.character(Idents(int5)),
                            check.names = F)

## We create a data frame with the Viper score per cell and its clusters
viper_scores_clusters <- viper_scores_df  %>%
  data.frame() %>% 
  rownames_to_column("cell") %>%
  gather(tf, activity, -cell) %>%
  inner_join(CellsClusters)

## We summarize the Viper scores by cellpopulation
summarized_viper_scores <- viper_scores_clusters %>% 
  group_by(tf, cell_type) %>%
  summarise(avg = mean(activity),
            std = sd(activity))
```

```{r}
head(summarized_viper_scores)
```
```{r}
## We select the 20 most variable TFs. (20*8 populations = 160)
highly_variable_tfs <- summarized_viper_scores %>%
  group_by(tf) %>%
  mutate(var = var(avg))  %>%
  ungroup() %>%
  top_n(160, var) %>%
  distinct(tf)

## We prepare the data for the plot
summarized_viper_scores_df <- summarized_viper_scores %>%
  semi_join(highly_variable_tfs, by = "tf") %>%
  dplyr::select(-std) %>%   
  spread(tf, avg) %>%
  data.frame(row.names = 1, check.names = FALSE) 
```

# Heatmap of DoRothEA 20 most variable TF activities per cluster
```{r, fig.height=5, fig.width=5}
palette_length = 100
my_color = colorRampPalette(c("Darkblue", "white","red"))(palette_length)

my_breaks <- c(seq(min(summarized_viper_scores_df), 0, 
                   length.out=ceiling(palette_length/2) + 1),
               seq(max(summarized_viper_scores_df)/palette_length, 
                   max(summarized_viper_scores_df), 
                   length.out=floor(palette_length/2)))

viper_hmap <- pheatmap(t(summarized_viper_scores_df),fontsize=14, 
                       fontsize_row = 10, 
                       color=my_color, breaks = my_breaks, 
                       main = "DoRothEA (ABCDEF)", angle_col = 90,
                       treeheight_col = 0,  border_color = NA) 
```
```{r}
saveRDS(summarized_viper_scores_df, file = "MarkersandPlots/Dor_int5_Mes_summarized_viper_scores_df.RDS")
```


```{r}
## We select the top5 TF per cluster
top5.tf <- All.markers.do %>% 
  group_by(cluster) %>% 
  top_n(n=5, wt = avg_log2FC) %>%
  rename(tf = gene) %>%
  distinct(tf)


## We prepare the data for the plot
summarized_viper_scores_top5_df <- summarized_viper_scores %>%
  semi_join(top5.tf, by = "tf") %>%
  dplyr::select(-std) %>%   
  spread(tf, avg) %>%
  data.frame(row.names = 1, check.names = FALSE) 
```

# Heatmap of DoRothEA top5 TF activities per DoRothEA cluster
```{r, fig.height=5, fig.width=5}
palette_length = 100
my_color = colorRampPalette(c("Darkblue", "white","red"))(palette_length)

my_breaks_top5 <- c(seq(min(summarized_viper_scores_top5_df), 0, 
                   length.out=ceiling(palette_length/2) + 1),
               seq(max(summarized_viper_scores_top5_df)/palette_length, 
                   max(summarized_viper_scores_top5_df), 
                   length.out=floor(palette_length/2)))

viper_hmap_top5 <- pheatmap(t(summarized_viper_scores_top5_df),fontsize=14, 
                       fontsize_row = 10, 
                       color=my_color, breaks = my_breaks, 
                       main = "DoRothEA (ABCDEF)", angle_col = 90,
                       treeheight_col = 0,  border_color = NA) 
```
```{r}
saveRDS(summarized_viper_scores_top5_df, file = "MarkersandPlots/Dor_int5_Mes_summarized_viper_scores_top5_df.RDS")
```

```{r}
## We create a data frame containing the cells and their clusters
Idents(int5) <- "Clusters_KOtype"
CellsClusters_seurat <- data.frame(cell = names(Idents(int5)), 
                            cell_type = as.character(Idents(int5)),
                            check.names = F)

## We create a data frame with the Viper score per cell and its clusters
viper_scores_clusters_seurat <- viper_scores_df  %>%
  data.frame() %>% 
  rownames_to_column("cell") %>%
  gather(tf, activity, -cell) %>%
  inner_join(CellsClusters_seurat)

## We summarize the Viper scores by cellpopulation
summarized_viper_scores_seurat <- viper_scores_clusters_seurat %>% 
  group_by(tf, cell_type) %>%
  summarise(avg = mean(activity),
            std = sd(activity))

## We select the top5 TF per cluster
top5.sc.tf <- All.markers %>% 
  group_by(cluster) %>% 
  top_n(n=5, wt = avg_log2FC) %>%
  rename(tf = gene) %>%
  distinct(tf)


## We prepare the data for the plot
summarized_viper_scores_top5sc_df <- summarized_viper_scores_seurat %>%
  semi_join(top5.sc.tf, by = "tf") %>%
  dplyr::select(-std) %>%   
  spread(tf, avg) %>%
  data.frame(row.names = 1, check.names = FALSE) 
```

# Heatmap of DoRothEA top5 TF activities per Seurat cluster by KO_type
```{r, fig.height=5, fig.width=5}
palette_length = 100
my_color = colorRampPalette(c("Darkblue", "white","red"))(palette_length)

my_breaks_top5sc <- c(seq(min(summarized_viper_scores_top5sc_df), 0, 
                   length.out=ceiling(palette_length/2) + 1),
               seq(max(summarized_viper_scores_top5sc_df)/palette_length, 
                   max(summarized_viper_scores_top5sc_df), 
                   length.out=floor(palette_length/2)))

viper_hmap_top5sc <- pheatmap(t(summarized_viper_scores_top5sc_df),fontsize=14, 
                       fontsize_row = 10, 
                       color=my_color, breaks = my_breaks_top5sc, 
                       main = "DoRothEA (ABCDEF)", angle_col = 90,
                       treeheight_col = 0,  border_color = NA) 
```
```{r}
saveRDS(summarized_viper_scores_top5sc_df, file = "MarkersandPlots/Dor_int5_Mes_summarized_viper_scores_top5sc_df.RDS")
```

# Heatmap of DoRothEA top20 TF activities WT vs KO for each seurat cluster
```{r}
DefaultAssay(object = int5) <- "dorothea"
Idents(int5) <- "Clusters_KOtype"
palette_length = 100
my_color = colorRampPalette(c("Darkblue", "white","red"))(palette_length)

for(cluster in unique(int5$Clusters)){
  ident1 <- paste(cluster,"WT",sep = "_")
  ident2 <- paste(cluster,"KO",sep = "_")
  markers <- FindMarkers(int5, ident.1 = ident1, ident.2 = ident2, verbose = FALSE)
  top20 <- head(markers,20) %>%
    add_rownames(var = "tf") %>%
    distinct(tf)
  summarized_viper_scores_cluster <- summarized_viper_scores_seurat %>%
    filter(cell_type %in% c(ident1, ident2)) %>%
    semi_join(top20, by = "tf") %>%
    dplyr::select(-std) %>%   
    spread(tf, avg) %>%
    data.frame(row.names = 1, check.names = FALSE) 
  my_breaks_cluster <- c(seq(min(summarized_viper_scores_cluster), 0, 
                             length.out=ceiling(palette_length/2) + 1),
                         seq(max(summarized_viper_scores_cluster)/palette_length, 
                             max(summarized_viper_scores_cluster), 
                             length.out=floor(palette_length/2)))
  
  viper_hmap_cluster <- pheatmap(t(summarized_viper_scores_cluster),fontsize=14, 
                                 fontsize_row = 10, 
                                 color=my_color, breaks = my_breaks_cluster, 
                                 main = paste("DoRothEA",cluster,"WTvsKO"), angle_col = 90,
                                 treeheight_col = 0,  border_color = NA)
  
  saveRDS(summarized_viper_scores_cluster, 
          file = paste0("MarkersandPlots/Dor_int5_Mes_summarized_viper_scores_cluster_",cluster,"_WTvsKO.RDS"))

}

```

# Heatmap of DoRothEA top20 TF activities WT vs KO for each DoRothEA cluster

```{r}
int5$dorothea_KOtype <- paste(int5$dorothea_snn_res.0.5, int5$KO_type, sep = "_")
Idents(int5) <- "dorothea_KOtype"

CellsClusters_doko <- data.frame(cell = names(Idents(int5)), 
                            cell_type = as.character(Idents(int5)),
                            check.names = F)

## We create a data frame with the Viper score per cell and its clusters
viper_scores_clusters_doko <- viper_scores_df  %>%
  data.frame() %>% 
  rownames_to_column("cell") %>%
  gather(tf, activity, -cell) %>%
  inner_join(CellsClusters_doko)

## We summarize the Viper scores by cellpopulation
summarized_viper_scores_doko <- viper_scores_clusters_doko %>% 
  group_by(tf, cell_type) %>%
  summarise(avg = mean(activity),
            std = sd(activity))
```

```{r}
DefaultAssay(object = int5) <- "dorothea"
Idents(int5) <- "dorothea_KOtype"
palette_length = 100
my_color = colorRampPalette(c("Darkblue", "white","red"))(palette_length)

for(cluster in unique(int5$dorothea_snn_res.0.5)){
  Idents(int5) <- "dorothea_KOtype"
  ident1 <- paste(cluster,"WT",sep = "_")
  ident2 <- paste(cluster,"KO",sep = "_")
  markers <- FindMarkers(int5, ident.1 = ident1, ident.2 = ident2, verbose = FALSE)
  top20 <- head(markers,20) %>%
    add_rownames(var = "tf") %>%
    distinct(tf)
  summarized_viper_scores_cluster <- summarized_viper_scores_doko %>%
    filter(cell_type %in% c(ident1, ident2)) %>%
    semi_join(top20, by = "tf") %>%
    dplyr::select(-std) %>%   
    spread(tf, avg) %>%
    data.frame(row.names = 1, check.names = FALSE) 
  my_breaks_cluster <- c(seq(min(summarized_viper_scores_cluster), 0, 
                             length.out=ceiling(palette_length/2) + 1),
                         seq(max(summarized_viper_scores_cluster)/palette_length, 
                             max(summarized_viper_scores_cluster), 
                             length.out=floor(palette_length/2)))
  
  viper_hmap_cluster <- pheatmap(t(summarized_viper_scores_cluster),fontsize=14, 
                                 fontsize_row = 10, 
                                 color=my_color, breaks = my_breaks_cluster, 
                                 main = paste("DoRothEA dorthoea cluster",cluster,"WTvsKO"), angle_col = 90,
                                 treeheight_col = 0,  border_color = NA)
  
  saveRDS(summarized_viper_scores_cluster, 
          file = paste0("MarkersandPlots/Dor_int5_Mes_summarized_viper_scores_cluster_Dor",cluster,"_WTvsKO.RDS"))
  
}
```

Note dorothea cluster 7 has no WT cells
```{r}
table(int5$dorothea_KOtype)
```


```{r}
saveRDS(int5, file = "RDSfiles/int5.Mes.subset.dorothea.RDS")
```

```{r}
sessionInfo()
```

